---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
getSymbols("MSFT", from = "2011-07-08" , to = "2011-07-19")
```

# CandleStickPattern

This package finds common candle stick patterns using daily data (OHLC data).

## Overview

Currently, it covers the following patterns:

- Doji family (1-day pattern)
    - Doji
    - Dragonfly Doji
    - Gravestone Doji

- Harmer family (2-day pattern)
    - Harmer
    - Hangman
    - Inverted Harmer
    - Shooting Star

- Engulfing family  (2-day pattern)
    - Bullish Engulfing
    - Bearish Engulfing
    
- Harami family (2-day pattern)
    - Bullish Harami
    - Bearish Harami
    
- Reversal family (2-day pattern)
    - Pierceing line
    - Dark cloud Cover
    
- Kicking family  (2-day pattern)
    - Kicking up
    - Kicking down

- Three in a row (3-day pattern)
    - Three white soldiers
    - Three Black crows
    
- Star (3-day pattern)
    - Morning star
    - Evening star
    
- Three Method (5-day pattern)
    - Rising three
    - Falling three
    
## Installation

To install this package, it is the easist to use devtools:

```{r, eval = FALSE}
#install.packages("devtools")
library(devtools)
```

Then we can use the `install_github` to get the package directly:

```{r, eval = FALSE}
#install_github("kochiuyu/CandleStickPattern")
library(CandleStickPattern)
```


## Usage

The following demonstrates the usage of the pacakge using Microsoft (ticker: MSFT). We use **quantmod** package to download data:

```{r, eval=FALSE}
library(quantmod)
getSymbols("MSFT", from = "2011-07-08" , to = "2011-07-19")
```

We first plot the candle chart first: 

```{r}
candleChart(MSFT,theme='white')
```

We can see that July 11, 12, 13 and 18 are likely to follow **doji** pattern. Let us see if we can capture using the function `doji`:

```{r}
library(CandleStickPattern)
doji(MSFT)
```
