---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
getSymbols("MSFT", from = "2011-07-08" , to = "2011-07-19")
```

# CandleStickPattern

This package finds common candle stick patterns using daily data (OHLC data).

## Overview

The current version package covers the following patterns:

- Doji family (1-day pattern)
    - Doji `doji()`
    - Dragonfly Doji `dragonfly.doji()`
    - Gravestone Doji `gravestone.doji()`

- Hammer family (2-day pattern)
    - Hammer/Hangman `hammer()`
    - Inverted Hammer/Shooting Star `inverted.hammer`

- Engulfing family  (2-day pattern)
    - Bullish Engulfing `bullish.engulf()`
    - Bearish Engulfing `bearish.engulf()`
    
- Harami family (2-day pattern)
    - Bullish Harami  `bullish.engulf()`
    - Bearish Harami  `bearish.harami()`
    
- Reversal family (2-day pattern)
    - Piercing line  `piercing.line()`
    - Dark cloud Cover `dark.cloud.cover()`
    
- Kicking family  (2-day pattern)
    - Kick up   `kick.up()`
    - Kick down  `kick.down()`

- Three-in-a-row (3-day pattern)
    - Three white soldiers `three.white.soliders()`
    - Three Black crows    `three.black.crows()`
    
- Star (3-day pattern)
    - Morning star  `morning.star()`
    - Evening star  `evening.star()`
    
- Three Method (5-day pattern)
    - Rising three   `rising.three()`
    - Falling three  `falling.three()`

Moreover, it captures trends using exponential moving average (EMA):

- EMA trends
    - up trend    `up.trend()`
    - down trend  `down.trend()`

## Installation

To install this package, it is the easist to install `devtools` package first:

```{r, eval = FALSE}
install.packages("devtools")
```

Now load the package using `library()`
```{r, eval = FALSE}
library(devtools)
```

Then we can use the `install_github` to get the package directly:

```{r, eval = FALSE}
install_github("kochiuyu/CandleStickPattern")
```

Finally, we load the package using `library()`

```{r, eval = FALSE}
library(CandleStickPattern)
```

## Usage

You can see the details of how each function works in [vignette](http://htmlpreview.github.io/?https://github.com/kochiuyu/CandleStickPattern/blob/master/vignettes/CandleStick.html)

The following demonstrates the usage of the function `doji()` using Microsoft (ticker: MSFT). We use **quantmod** package to download data:

```{r, eval=FALSE}
library(quantmod)
getSymbols("MSFT", from = "2011-07-08" , to = "2011-07-19")
```

We first plot the candle chart first: 

```{r}
candleChart(MSFT,theme='white')
```

We can see that July 11, 12, 13, and 18 are likely to follow **doji** pattern. Let us see if we can capture using the function `doji`:

```{r}
library(CandleStickPattern)
doji(MSFT)
```
